### 2020-05-11 writeup/proof_transfer_ex.tex
sec 3.1

\section{Supplementary Materials for Section \ref{sec_transfer}}

As a warm up, we show that when $\beta_s = \beta_t$, then the transfer is always positive.

\begin{proposition}\label{prop_monotone}
	Suppose that $n > p$.
  When there is no model shift, i.e. $\beta_s = \beta_t$, adding the source task data always reduces the estimation error and the test error for the target task, i.e.
	\begin{align}
%		\err(\hat{\beta}_{t}^{\MTL})  &\le \err(\hat{\beta}_t^{\STL}), \text{ and} \label{eq_mono_e}\\
		\te(\hat{\beta}_{t}^{\MTL}) &\le \te(\hat{\beta}_t^{\STL}). \label{eq_mono_te}
	\end{align}
\end{proposition}

\begin{proof}
%	Equation \eqref{eq_mono_e} is simply because
%		\[ \err(\hat{\beta}_{s,t}) = \bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}} \le \bigtr{(X_1^{\top}X_1)^{-1}} = \err(\hat{\beta}_t). \]
%	Equation \eqref{eq_mono_te} follows because
	Recall that $\hat{\beta}_t^{\MTL} = \hat{w} \cdot \hat{B}$.
	By the optimality of $\hat{w}$, we have by setting $w = 1$ in equation \eqref{eq_te_mtl}
	\begin{align*}
		\te(\hat{\beta}_t^{\MTL}) &\le \sigma^2 \cdot \bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} \\
		&= \sigma^2 \cdot \bigtr{\bigbrace{\Sigma_2^{-1/2}X_1^{\top}X_1\Sigma_2^{-1/2} + \Sigma_2^{-1/2}X_2^{\top}X_2\Sigma^{-1/2}}^{-1}} \\
		&\le \bigtr{\bigbrace{\Sigma_2^{-1/2}X_2^{\top}X_2\Sigma_2^{-1/2}}^{-1}}
			= \bigtr{(X_2^{\top}X_2)^{-1} \Sigma_2} = \te(\hat{\beta}_t^{\STL}),
	\end{align*}
	which concludes the proof.
\end{proof}

As a remark, we can derive a similar result for the estimation error as well. The details are omitted.

>>>>

\section{Supplementary Materials for Section \ref{sec_insight}}

From \cite{WZR20}, we know that we need to explicitly restrict the capacity $r$ of $B$ so that there is transfer between the two tasks.
for the rest of the section, we shall consider the case when $r=1$ we are considering the case of two tasks.
Here, equation \eqref{eq_mtl} simplifies to the following
\[ f(B; w_1, w_2) = \bignorm{X_1 B w_1 - Y_1}^2 + \bignorm{X_2 B w_2 - Y_2}^2, \]
where $B\in\real^p$ and $w_1, w_2$ are both real numbers.
To solve the above, suppose that $w_1, w_2$ are fixed, by local optimality, we solve $B$ as
\begin{align*}
	\hat{B}(w) &= (w_1^2 X_1^{\top}X_1 + w_2^2 X_2^{\top}X_2)^{-1} (w_1 X_1^{\top}Y_1 + w_2 X_2^{\top}Y_2) \\
	&= \frac{1}{w_2} (w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} (w X_1^{\top}Y_1 + X_2^{\top}Y_2) \\
	&= \frac{1}{w_2}\bigbrace{\beta_t + (w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\bigbrace{X_1^{\top}X_1(w\beta_s - w^2\beta_t) + (w X_1^{\top}\varepsilon_1 + X_2^{\top}\varepsilon_2)}},
\end{align*}
where we denote $w = w_1 / w_2$.
As a remark, when $w = 1$, we recover the linear regression estimator.

\medskip
\noindent\textit{Jointly optimizing over both tasks.}
Using a validation set that is sub-sampled from the original training dataset, we get a validation loss as follows
\begin{align}
		&\val(\hat{B}; w_1, w_2) \nonumber\\
	=& n_1 \cdot \bigbrace{\bignorm{\Sigma_1^{1/2}(w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_2^{\top}X_2(\beta_s - w\beta_t)}^2 + \sigma^2 \cdot \bigtr{(w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_1}} \nonumber \\
	+& n_2 \cdot \bigbrace{w^2\bignorm{\Sigma_2^{1/2}(w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1(\beta_s - w\beta_t)}^2 + \sigma^2 \cdot \bigtr{(w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}} \label{eq_val_mtl}
\end{align}
Minimizing the above loss is akin to performing multi-task training in practice.
Let $\hat{w}$ denote the minimizer of $\val(\hat{B}; w_1, w_2)$ over $w\in\real$.
We will denote $\hat{\beta}_t^{\MTL} = w_{2}\hat{B}(\hat{w})$.

Our goal is to study under model and covariate shifts, whether multi-task learning helps learn the target task better than single-task learning.
The baseline where we solve the target task with its own data is
\begin{align*}
	te(\hat{\beta}_t^{\STL}) = \sigma^2 \cdot \bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}}, \text{ where } \hat{\beta}_t^{\STL} = (X_2^{\top}X_2)^{-1} X_2^{\top}Y_2.
\end{align*}
Clearly, whether $\hat{\beta}_t^{\MTL}$ outperforms $\hat{\beta}_t^{\STL}$ depends on the covariate matrices, the difference of the task models, and the number of per-task data points.

\begin{lemma}\label{lem_minv}[\todo{ref?}]
	In the setting when $n_2 = c_2 p$ we have that as $p$ goes to infinity,
	\[ \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2} = \frac{p}{n_2 - p}. \]
\end{lemma}

\begin{lemma}\label{lem_cov_shift}
	In the setting of Theorem \ref{thm_model_shift}, we have with high probability $1-{\rm o}(1)$,
\begin{align}\label{lem_cov_shift_eq}
\tr ((w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2) = \frac{1}{n_1+n_2}\cdot \tr \left( \frac{1}{a_1 M^\top M + a_2} \right) +\bigo{ n^{-1/2+\epsilon}},
\end{align}
	for any constant $\epsilon>0$.
\end{lemma}
We will give the proof of this lemma in Section \ref{sec_maintools}.

\subsection{Proof of Proposition \ref{prop_transition}}

Hence the task models have distance $d^2\cdot p$ in expectation.
	We first consider $\Sigma_2 = \id$. In this case, we can simplify $\Delta_{\beta}$ as follows
	\begin{align} \label{eq_delta_simple}
		\Delta_{\beta} \define d^2 \cdot \sum_{i=1}^p \frac{(1 + a_3)\lambda_i^2 + a_4 \lambda_i^4}{(a_1 \lambda_i^2 + a_2)^2}.
	\end{align}
	Now we solve the equations \eqref{eq_a2}, \eqref{eq_a3}, \eqref{eq_a4} to get
	\begin{align}
		a_1 = \frac{c_1(c_1 + c_2 - 1)}{(c_1 + c_2)^2},
		a_2 = \frac{c_2(c_1 + c_2 - 1)}{(c_1 + c_2)^2},
		a_3 = \frac{c_2}{(c_1 + c_2)(c_1 + c_2 - 1)},
		a_4 = \frac{c_1}{(c_1 + c_2)(c_1 + c_2 - 1)}.
	\end{align}
%{\color{blue}if $\Sigma_1=\Sigma_2=\id$, then
%	\begin{align}
%		a_1 = c_1 \left( 1- \gamma_n\right) , \quad
%		a_2 = c_2 \left( 1- \gamma_n\right), \quad
%		a_3 = \frac{\gamma_n c_2}{1-\gamma_n}, \quad
%		a_4 =  \frac{\gamma_n c_1}{1-\gamma_n}.
%	\end{align}
%	where $\gamma_n=p/n$, $c_1=n_1/n$, and $c_2=n_2/n$.
%}

	Then we obtain
	\begin{align}
		\Delta_{\beta} = p \cdot d^2 \cdot \frac{c_1^2 (c_1 + c_2)}{(c_1 + c_2 - 1)^3},
		\Delta_{\vari} = \sigma^2 \cdot \frac{c_1}{(c_2 - 1)(c_1 + c_2 - 1)}.
	\end{align}

### 2020-05-11 writeup/transfer.tex
sec 3.1

\section{Illustrative Examples and Their Insights on Transfer}\label{sec_transfer}

@@

We describe examples based on Theorem \ref{thm_model_shift} to show several conceptual insights.

\textbf{Example.}
	We consider a simple setting where $\Sigma_1 = \id$.
	Suppose that $\beta_s - \beta_t$ is i.i.d. with mean $0$ and variance $d^2$. Hence the task models have distance $d^2\cdot p$ in expectation.

	We first consider $\Sigma_2 = \id$. In this case, we can simplify $\Delta_{\beta}$ as follows
	\begin{align} \label{eq_delta_simple}
		\Delta_{\beta} \define d^2 \cdot \sum_{i=1}^p \frac{(1 + a_3)\lambda_i^2 + a_4 \lambda_i^4}{(a_1 \lambda_i^2 + a_2)^2}.
	\end{align}
	Now we solve the equations \eqref{eq_a2}, \eqref{eq_a3}, \eqref{eq_a4} to get
	\begin{align}
		a_1 = \frac{c_1(c_1 + c_2 - 1)}{(c_1 + c_2)^2},
		a_2 = \frac{c_2(c_1 + c_2 - 1)}{(c_1 + c_2)^2},
		a_3 = \frac{c_2}{(c_1 + c_2)(c_1 + c_2 - 1)},
		a_4 = \frac{c_1}{(c_1 + c_2)(c_1 + c_2 - 1)}.
	\end{align}
%{\color{blue}if $\Sigma_1=\Sigma_2=\id$, then
%	\begin{align}
%		a_1 = c_1 \left( 1- \gamma_n\right) , \quad
%		a_2 = c_2 \left( 1- \gamma_n\right), \quad
%		a_3 = \frac{\gamma_n c_2}{1-\gamma_n}, \quad
%		a_4 =  \frac{\gamma_n c_1}{1-\gamma_n}.
%	\end{align}
%	where $\gamma_n=p/n$, $c_1=n_1/n$, and $c_2=n_2/n$.
%}

	Then we obtain
	\begin{align}
		\Delta_{\beta} = p \cdot d^2 \cdot \frac{c_1^2 (c_1 + c_2)}{(c_1 + c_2 - 1)^3},
		\Delta_{\vari} = \sigma^2 \cdot \frac{c_1}{(c_2 - 1)(c_1 + c_2 - 1)}.
	\end{align}

	We demonstrate our result with a simulation. (\todo {uses the tighter bound Proposition \ref{prop_model_shift_tight}?})
	We consider a setting where $p = 200$, $n_1 = 90p$, $n_2 = 30p$.
	\todo{Fill in other params.}
	We fix the target task and vary the source task, by varying the task model distance parameter $d$.
	We show that Theorem \ref{thm_model_shift} predicts whether we can get positive or negative transfer.
	Figure \ref{fig_model_shift_phasetrans} shows the result.
	We obtain the following insight from the simulation.
	\begin{itemize}
		\item Adding the source task has the effect of reducing the variance of the estimator, independent of the model shift.
		\item Model shift introduces an additional bias term, which scales with $d^2$, the distance of the two task models.
		Hence, the type of transfer is determined by the tradeoff between the bias caused by model shift and the reduction of variance.
	\end{itemize}

>>>>

\section{Illustrative Examples and Their Insights on Transfer}\label{sec_insight}

@@

We compare the test error of $\hat{\beta}_t^{\MTL}$ to that of $\hat{\beta}_t^{\STL}$.
For a simple example, we show that whether $\hat{\beta}_t^{\MTL}$ performs better than $\hat{\beta}_t^{\STL}$ is determined by the distance of the task models.
We derive a sharp threshold when positive transfer transitions to negative transfer, as a ratio between the model distance and the noise level.

\textbf{Example.}
	Consider a setting where $\Sigma_1 = \Sigma_2 = \id$, in other words there is no covariate shift between the two tasks.
	For the task models, suppose that $\beta_t$ has i.i.d. entries with mean zero and variance $\kappa^2$ and $\beta_s - \beta_t$ also has i.i.d. entries with mean $0$ and variance $d^2$.
	We have $n = c \cdot p$ data points from each task.

	We illustrate the example in a synthetic setting.
	We demonstrate our result with a simulation. (\todo {uses the tighter bound Proposition \ref{prop_model_shift_tight}?})
	We consider a setting where $p = 200$, $n_1 = 90p$, $n_2 = 30p$.
	\todo{Fill in other params.}
	We fix the target task and vary the source task, by varying the task model distance parameter $d$.
	We show that Theorem \ref{thm_model_shift} predicts whether we can get positive or negative transfer.
	Figure \ref{fig_model_shift_phasetrans} shows the result.

Specifically, the transition threshold is derived in the following proposition.
\begin{proposition}\label{prop_transition}
	In the example described above, we have the following deterministic conditions to determine whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$:
	\begin{itemize}
		\item If ${p\cdot d^2}/{\sigma^2} < \frac{2(2c-1)^2}{(c-1)(2c+1)} - o(1)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If ${p\cdot d^2}/{\sigma^2} \ge \frac{2(2c-1)^2}{(c-1)(2c+1)} + o(1)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{proposition}

We obtain the following insight from Proposition \ref{prop_transition}.
First, adding the source task has the effect of reducing the variance of the estimator, independent of the model shift.
On the other hand, model shift introduces an additional bias term, which scales with $d^2$, the distance of the two task models.
Hence, the type of transfer is determined by the trade-off between the bias caused by model shift and the reduction of variance.
